### 2015-07-07 narrative/qos.tex
minor changes.

\subsubsection{Resource Management}
\paragraph{Background:} There is an important connection between finding the resources
that can fulfill a particular request and the performance predictability of that request, and this demands
careful resource management strategies that can deal with placing and retrieving a large number of data 
chunks spread across the entire storage hierarchy. 
%This connection is particularly important for data retrieval 
%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
%component of overall read latency. 
Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
scalability, because they require a single entity that tracks updates in the system. 
Traditional file systems are good examples for this - most updates to files require updates 
to their centrally maintained block allocation tables. 
The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
resources at request time. Pile systems allow very efficient writes because clients can write whereever
there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
is possible, e.g., when it is unclear whether the resource actually exists. 
Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
partitioning of the search space and a method to limit the search to a small number of partitions.

@@

\subsubsection{QoS}
\paragraph{Background:} Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest.
This can be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

@@

\paragraph{Challenges:} The deep and heterogenous memory and storage
hierarchy we are assuming for this proposal complicates the
relationship between workload-dependent and workload-independent
metrics: the performance of a task can significantly differ depending
on what levels of the hierarchy are involved. Furthermore, an
important goal of the proposed project is to enable applications
to reason about the trade-off between resolution and latency, adding
yet another dimension to how tasks are mapped to resources.

@@

\item The complexity of a utilization model involving the entire
storage hieararchy with 100,000s of devices is potentially daunting.
However, the utilization model can be simplified by modeling classes
of devices as well as classes of requests. In particular, each
device could restrict access to its content via a set of well-defined
methods with known, absolute performance properties.

>>>>

\subsubsection{Resource Management}
\paragraph{Background:} There is an important connection between finding the resources
that can fulfill a particular request and the performance predictability of that request, and this demands
careful resource management strategies that can deal with placing and retrieving a large number of data 
chunks spread across the entire storage hierarchy. 
%This connection is particularly important for data retrieval 
%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
%component of overall read latency. 
Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
scalability, because they require a single entity that tracks updates in the system. 
Traditional file systems are good examples for this - most updates to files require updates 
to their centrally maintained block allocation tables. 
The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
resources at request time. Pile systems allow very efficient writes because clients can write wherever
there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
is possible, e.g., when it is unclear whether the resource actually exists. 
Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
partitioning of the search space and a method to limit the search to a small number of partitions.

@@

\subsubsection{QoS}
\paragraph{Background:} Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to interference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
In the cases where an application arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest.
This can be automatically enforced via user- (or administrator-) specified \textit{policies}
(section \ref{}).

@@

\paragraph{Challenges:} The deep and heterogeneous memory and storage
hierarchy we are assuming for this proposal complicates the
relationship between workload-dependent and workload-independent
metrics: the performance of a task can significantly differ depending
on what levels of the hierarchy are involved. Furthermore, an
important goal of the proposed project is to enable applications
to reason about the trade-off between resolution and latency, adding
yet another dimension to how tasks are mapped to resources.

@@

\item The complexity of a utilization model involving the entire
storage hierarchy with 100,000s of devices is potentially daunting.
However, the utilization model can be simplified by modeling classes
of devices as well as classes of requests. In particular, each
device could restrict access to its content via a set of well-defined
methods with known, absolute performance properties.
